State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-04-14 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
paper updates

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
	f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \varsigma \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}.
\end{equation}
%Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. 
By substituting \ref{RateBasedInteractions} into \ref{SpikesToPotential}  we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
	v\left(\mathbf{r},t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega   w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'.
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
	Dh\left( t \right) = \delta \left( t \right),
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}\label{FinalFormContinuous}
	\frac{dv\left( \mathbf{r},t \right)}{dt} + \zeta v\left( \mathbf{r},t \right) = \int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)dr'} + u\left(\mathbf{r},t\right).
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model we discretize time using a first-order Euler method (see Appendix~\ref{Time Discretization}) giving
\begin{equation}\label{DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right),
\end{equation} 
where $T$ is the time step, $\xi = 1-\zeta T$ and $e_t\left(\mathbf{r}\right)$ is an iid spatially coloured, temporally white disturbance term with the covariance 
\begin{equation}
    cov(\mathbf{e}_t,\mathbf{e}_{t+\tau}) = \left\{ \begin{array}{*{20}{c}}
    \gamma\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
    0 & \mathrm{otherwise}. \\
    \end{array} \right.
\end{equation}
This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t,
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}.
\end{equation}

@@

Similar considerations need to be made regarding the representation of the dynamic field $w_t(\mathbf{r})$ by its basis functions decompositions. The minimum distance between adjacent basis functions also needs to satisfy Shannon's sampling theorem. Thus, the minimum distance between basis must satisfy
\begin{equation}
\Delta_b < \frac{1}{2\nu_c}.
\end{equation}
For the Gaussian basis functions $\phi(\mathbf{r})$ being considered, the basis width parameter $\sigma^2$ can also be inferred from spectral considerations \cite{Sanner1992,Scerri2009}. The Fourier transform of a Gaussian is another Gaussian with width $\sigma^2_{\nu}$ given by
\begin{equation}
\sigma^2_{\nu} = \frac{1}{\pi^2\sigma^2}.
\end{equation}
Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with the 3~dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma^2$.

@@

Due to the high dimensionality of the brain and the current electrode systems, we can not expect to have more sensors than basis functions.
\section{State and Parameter Estimation}
The estimation process is a two part iterative algorithm that yields the estimate of the connectivity kernel parameters and the posterior distribution of the field weights. In the parameter estimation step, Least Squares (LS) method is used which minimizes the sum of the squares of the errors on each state equation (see Appendix). In the state estimation step, given an estimate of $\boldsymbol \theta$ and $\xi$ the Unscented Rauch Tung Striebel Smoother (URTSS) \cite{Sarkka2010} was used to determine the posterior state sequence.
The (URTSS) uses the Unscented Kalman Filter (UKF) in the forward iteration \cite{Julier1997} along with a separate backward smoothing path to compute the smooth estimations. The method is appropriate for state estimation of stochastic nonlinear dynamical systems. Alternatives to the UKF are the Extended Kalman Filter (EKF) \cite{Haykin2001}  and Sequential Monte Carlo (SMC) filtering \cite{doucet2001}. The extended Kalman filter approximates the state transition equation by linearizing about the fixed points (by calculating the Jacobian) at each time step. The linearization maintains the Gaussianality in the model. The unscented Kalman filter approximates the \emph{a posteriori} state density by a Gaussian distribution using a minimal set of carefully chosen sigma points, whilst maintaining the nonlinearity in the system. This has been shown to give superior performance over the EKF in state estimation, as the EKF maintains a first order approximation where the UKF provides an approximation accurate at least to the second order. In addition, the computation of the Jacobian used in the
EKF can be problematic. SMC filtering can theoretically provide an exact \emph{a posteriori} state density for nonlinear systems. However, currently this method is not appropriate for our problem, due to computation demands.

The approximation of the state density is made by propagating the so called sigma points through the state-space model. The sigma points describe the mean and covariance of the state estimate. The number of sigma points needed is $2\mathbf L+1$ where $\mathbf L$ is the dimension of the stae space. The  columns of the sigma point matrix and their corresponding weights used in the UKF can be calculated as
\begin{equation}
 \mathcal X_{0}=\bar x
\end{equation}
\begin{equation}
 \mathcal X_{i}=\bar x+(\sqrt{(\mathbf L + \lambda)\mathbf P_x})_i \quad i=1, \dots, \mathbf L
\end{equation}
\begin{equation}
 \mathcal X_{i}=\bar x-(\sqrt{(\mathbf L + \lambda)\mathbf P_x})_{i-\mathbf L} \quad i=\mathbf L+1, \dots, 2\mathbf L
\end{equation}
\begin{equation}
 \mathbf W_0^{(m)}=\frac{\lambda}{\mathbf L+\lambda}
\end{equation}
\begin{equation}
 \mathbf W_0^{(c)}=\frac{\lambda}{\mathbf L+\lambda}+(1-\alpha^2+\beta)
\end{equation}
\begin{equation}
 \mathbf W_i^{(m)}=\mathbf W_i^{(c)}=\frac{1}{2(\mathbf L+\lambda)} \quad i=1, \dots, \mathbf 2L
\end{equation}
where the sacling parameter $\lambda$ is defined as
\begin{equation}
\lambda=\alpha^2(\mathbf L+\kappa)-\mathbf L 
\end{equation}
parameters $\alpha$, $\beta$ and $\kappa$ can be used to tune the filtering process \cite{Haykin2001}

@@

 The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t)\boldsymbol{\theta} - \xi\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\xi$ are the systems parameters. In forwards iteration since the observation equation is linear in parameters we only need to use Unsented transformation in prediction step. The update step equations remains the same as the standard Kalman filtering formulations. In backward smoothing path, estimation from the forward filtering is used to form the new set of sigma points in a similar way to the filterin step, however the predicted cross-covariance of the states need to be calculated. A summary of the algorithm is given in Algorithm 
\begin{algorithm}
\caption{The Unscented Kalman Filter}\label{CPTEMDAlgorithm}
\begin{algorithmic}[1]
\State Initialise
\begin{equation*}
\hat{\mathbf x}_0, \mathbf P_0
\end{equation*}
\State Forward iteration
For $t \in \left\lbrace 0,\cdots, N\right\rbrace $,
\begin{equation*}
\mathcal X_t^f=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^{f}-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
Propogate the sigma points through the dynamic model
\begin{equation}
\mathcal X_{i,t+1|t}^{f*}=Q(\mathcal X_t^f,u_t)
\end{equation}
\begin{equation*}
 \hat{\mathbf x}_{k +t}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1|t}^{f*}
\end{equation*}
\begin{equation*}
 \mathbf P_{t +1}^-=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e
\end{equation*}
\begin{equation*}
\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_v)^{-1}
\end{equation*}
\begin{equation*}
\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-})
\end{equation*}
\begin{equation*}
 \mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}
\end{equation*}
\State Backward iteration
\State initialisation
\begin{equation}
 \mathbf P_T^s= \mathbf P_T^f
\end{equation}
\begin{equation}
\hat{\mathbf x}^s_T= \hat{\mathbf x}^f_T
\end{equation}
For $t \in \left\lbrace N-1, \cdots, 0 \right\rbrace $
\begin{equation*}
\mathcal X_t^s=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^f-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
\end{algorithmic}
\end{algorithm}
\section{Results}

@@

\begin{eqnarray}\label{Poo}

@@

\begin{eqnarray}\label{Poo}

@@

@@

\begin{tabular} {c|c} 
	\hline\hline
	Symbol & Description \\ \hline
	$v(\mathbf{r},t)$ & mean membrane potential \\
	$g(\mathbf{r},t)$ & average action potential rate \\
	$\mathbf{r}$ & spatial location \\
	$t$ & time (s) \\
	$h(t)$ & post-synaptic response kernel \\
	$\eta(t)$ & Heaviside function \\
	$\zeta$ & inverse synaptic time constant \\
	$w(\mathbf{r},\mathbf{r}')$ & spatial connectivity kernel \\
	$f(\mathbf{r},t)$ & firing function rate \\
	$u(\mathbf{r},t)$ & external input \\
	$\Omega$ & spatial domain \\
	$f_{max}$ & maximal firing rate \\
	$\varsigma$ & slope of sigmoidal activation function \\
	$v_0$ & firing threshold \\
	$\delta(t)$ & Dirac-delta function \\
	$L$ & temporal differential operator \\
	$e(\mathbf{r},t)$ & field disturbance, covariance $\gamma$\\
	$m(\mathbf{r},\mathbf{r}')$ & sensor kernel, variance $\sigma_m^2$ \\
	$\varepsilon(\mathbf{r}_n,t)$ & observation noise, covariance $\Sigma_\epsilon$ \\
	$y(\mathbf{r}_n,t)$ & obseration \\
	$n$ & sensor index $n=1,...,N$ \\
	$T$ & time step \\
	$\xi$ & time constant parameter \\
	$\boldsymbol{\phi(r)}$ & vector of Gaussian basis functions \\
	$\mathbf{x}_t$ & state vector at time $t$ \\
	$\boldsymbol{\psi}$ & vector of connectivity kernel basis functions \\
	$\boldsymbol{\theta}$ & vector of connectivity kernel parameters \\
	$\boldsymbol{\Gamma}$ & inner product of field basis functions \\
	$q(.)$ & state transition function \\
	$k(.)$ & maps into to discrete state of input function $u(\mathbf{r},t)$ \\
	$\mathbf{e}_t$ & state disturbance, covariance $\Sigma_e$ \\
	$\mathbf{C}$ & observation matrix \\
	$V(\nu)$ & Spectrum of the dynamic field \\ 
	$\nu$ & Spatial frequency \\
	$\nu_c$ & Spatial cut-off frequency \\
\end{tabular}

\begin{tabular}{c|c}
	\hline\hline
	Symbol & Description \\ \hline
	$\Delta_s$ & Distance between adjacent sensors \\
	$\rho$ & Over-sampling parameter \\	
	$\Delta_b$ & Distance between field basis functions \\	
	$\sigma_{nu}^2$ & Variance of FT of Gaussian basis function \\
	$\sigma^2$ & Spatial variance of Gaussian in spatial domain \\
	$\hat{\mathbf{x}}$ & State estimate \\
	$\boldsymbol{\chi}$ & Matrix of sigma vectors \\
	$\hat{\mathbf{x}}_t^{f-}$ & Forward \emph{a priori} estimate \\
	$\hat{\mathbf{x}}_t^f$ & Forward \emph{a posteriori} estimate \\
	$\hat{\mathbf{x}}_t^{b-}$ & Backward \emph{a priori} estimate \\
	$\hat{\mathbf{x}}_t^{b}$ & Backward \emph{a posteriori} estimate \\
	$\mathbf{P}^f_t$ & Forward \emph{a posteriori} covariance matrix \\
	$\mathbf{P}^{f-}_t$ & Forward \emph{a priori} covariance matrix \\	
\end{tabular}
\section*{References}
\bibliographystyle{unsrt}
\bibliography{BrainIDE}

>>>>

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
	f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \varsigma \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}.
\end{equation}
%Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. 
By substituting \ref{RateBasedInteractions} into \ref{SpikesToPotential}  we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
	v\left(\mathbf{r},t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega   w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'.
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
	Dh\left( t \right) = \delta \left( t \right),
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
where $L=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}\label{FinalFormContinuous}
	\frac{dv\left( \mathbf{r},t \right)}{dt} + \zeta v\left( \mathbf{r},t \right) = \int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)dr'} + u\left(\mathbf{r},t\right).
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model we discretize time using a first-order Euler method (see Appendix~\ref{Time Discretization}) giving
\begin{equation}\label{DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right),
\end{equation} 
where $T$ is the time step, $\xi = 1-\zeta T$ and $e_t\left(\mathbf{r}\right)$ is an iid spatially coloured, temporally white disturbance term with the covariance 
\begin{equation}
    cov(\mathbf{e}_t,\mathbf{e}_{t+\tau}) = \left\{ \begin{array}{*{20}{c}}
    \gamma\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
    0 & \mathrm{otherwise}. \\
    \end{array} \right.
\end{equation}
This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t,
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}.
\end{equation}

@@

Similar considerations need to be made regarding the representation of the dynamic field $v_t(\mathbf{r})$ using the basis function decomposition. The minimum distance between adjacent basis functions also needs to satisfy Shannon's sampling theorem. Thus, the minimum distance between basis must satisfy
\begin{equation}
\Delta_b < \frac{1}{2\nu_c}.
\end{equation}
For the Gaussian basis functions $\phi(\mathbf{r})$ being considered, the basis width parameter $\sigma^2$ can also be inferred from spectral considerations \cite{Sanner1992,Scerri2009}. The Fourier transform of a Gaussian is another Gaussian with width $\sigma^2_{\nu}$ given by
\begin{equation}
\sigma^2_{\nu} = \frac{1}{\pi^2\sigma^2}.
\end{equation}
Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with the 3~dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma^2$.

@@

Due to the high dimensionality of the brain and the current electrode systems, we can not expect to have more sensors than basis functions.
\section{State and Parameter Estimation}
The estimation process is a two part iterative algorithm that yields the estimate of the connectivity kernel parameters and the posterior distribution of the field weights. In the parameter estimation step, Least Squares (LS) method is used which minimizes the sum of the squares of the errors on each state equation (see Appendix). In the state estimation step, given an estimate of $\boldsymbol \theta$ and $\xi$ the Unscented Rauch Tung Striebel Smoother (URTSS) \cite{Sarkka2010} was used to determine the posterior state sequence .
The (URTSS) uses the Unscented Kalman Filter (UKF) in the forward iteration \cite{Julier1997} to calculate a forward posterior estimate $\hat{\mathbf x}_t^{f} $ along with a separate backward smoothing path to compute the smooth estimations $\hat{\mathbf x}_t^{b} $. The method is appropriate for state estimation of stochastic nonlinear dynamical systems. Alternatives to the UKF are the Extended Kalman Filter (EKF) \cite{Haykin2001}  and Sequential Monte Carlo (SMC) filtering \cite{doucet2001}. The extended Kalman filter approximates the state transition equation by linearizing about the fixed points (by calculating the Jacobian) at each time step. The linearization maintains the Gaussianality in the model. The unscented Kalman filter approximates the \emph{a posteriori} state density by a Gaussian distribution using a minimal set of carefully chosen sigma points, whilst maintaining the nonlinearity in the system. This has been shown to give superior performance over the EKF in state estimation, as the EKF maintains a first order approximation where the UKF provides an approximation accurate at least to the second order. In addition, the computation of the Jacobian used in the
EKF can be problematic. SMC filtering can theoretically provide an exact \emph{a posteriori} state density for nonlinear systems. However, currently this method is not appropriate for our problem, due to computation demands.

The approximation of the state density is made by propagating the so called sigma points through the state-space model. The sigma points describe the mean and covariance of the state estimate. The number of sigma points needed is $2L+1$ where $ L$ is the dimension of the state space. The  columns of the sigma point matrix and their corresponding weights used in the UKF can be calculated as
\begin{equation}
 \mathcal X_{0}=\bar x
\end{equation}
\begin{equation}
 \mathcal X_{i}=\bar x+(\sqrt{( L + \lambda)\mathbf P_x})_i \quad i=1, \dots,  L
\end{equation}
\begin{equation}
 \mathcal X_{i}=\bar x-(\sqrt{( L + \lambda)\mathbf P_x})_{i- L} \quad i= L+1, \dots, 2 L
\end{equation}
\begin{equation}
 \mathbf W_0^{(m)}=\frac{\lambda}{ L+\lambda}
\end{equation}
\begin{equation}
 \mathbf W_0^{(c)}=\frac{\lambda}{ L+\lambda}+(1-\alpha^2+\beta)
\end{equation}
\begin{equation}
 \mathbf W_i^{(m)}=\mathbf W_i^{(c)}=\frac{1}{2( L+\lambda)} \quad i=1, \dots, 2L
\end{equation}
where the scaling parameter $\lambda$ is defined as
\begin{equation}
\lambda=\alpha^2( L+\kappa)- L 
\end{equation}
and parameters $\alpha$, $\beta$ and $\kappa$ are used to tune the filtering process \cite{Haykin2001}. It should be noted that $\bar x$ is $ \hat{\mathbf x}_t^{f} $ and $ \hat{\mathbf x}_t^{b} $ with their corresponding covariance matrices in filtering and smoothing iterations respectively.

@@

 The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t)\boldsymbol{\theta} - \xi\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\xi$ are the systems parameters. In forwards iteration since the observation equation is linear in parameters we only need to use Unsented transformation in prediction step. The update step equations remains the same as the standard Kalman filtering formulations. In backward smoothing path, estimation from the forward filtering is used to form the new set of sigma points in a similar way to the filterin step, however the predicted cross-covariance of the states need to be calculated. A summary of the algorithm is given in Algorithm 
\begin{algorithm}
\caption{The Unscented Kalman Filter}\label{CPTEMDAlgorithm}
\begin{algorithmic}[1]
\State Initialise
\begin{equation*}
\hat{\mathbf x}_0, \mathbf P_0
\end{equation*}
\State Forward iteration
For $t \in \left\lbrace 0,\cdots, N\right\rbrace $,
\begin{equation*}
\mathcal X_t^f=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^{f}-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
Propogate the sigma points through the state equation
\begin{equation}
\mathcal X_{i,t+1|t}^{f*}=Q(\mathcal X_t^f,u_t)
\end{equation}
\begin{equation*}
 \hat{\mathbf x}_{k +t}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1|t}^{f*}
\end{equation*}
\begin{equation*}
 \mathbf P_{t +1}^-=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e
\end{equation*}
\begin{equation*}
\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_v)^{-1}
\end{equation*}
\begin{equation*}
\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-})
\end{equation*}
\begin{equation*}
 \mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}
\end{equation*}
\State Backward iteration
\State initialisation
\begin{equation}
 \mathbf P_T^s= \mathbf P_T^f
\end{equation}
\begin{equation}
\hat{\mathbf x}^s_T= \hat{\mathbf x}^f_T
\end{equation}
For $t \in \left\lbrace T-1, \cdots, 0 \right\rbrace $
\begin{equation*}
\mathcal X_t^s=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^f-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
\State Propogate the sigma points through the state equation similar to the forward step and compute

\end{algorithmic}
\end{algorithm}
\section{Results}

@@

\begin{eqnarray}\label{ReducingState}

@@

\begin{eqnarray}\label{StartofReduction}

@@

Substituting Eq.~\ref{DefGamma} into Eq.~\ref{StartofReduction} and cross-multiplying by $\boldsymbol{\Gamma}^{-1}$ gives 

@@

\begin{tabular} {c|c} 
	\hline\hline
	Symbol & Description \\ \hline
	$v(\mathbf{r},t)$ & mean membrane potential \\
	$g(\mathbf{r},t)$ & average action potential rate \\
	$\mathbf{r}$ & spatial location \\
	$t$ & time (s) \\
	$h(t)$ & post-synaptic response kernel \\
	$\eta(t)$ & Heaviside function \\
	$\zeta$ & inverse synaptic time constant \\
	$w(\mathbf{r},\mathbf{r}')$ & spatial connectivity kernel \\
	$f(\mathbf{r},t)$ & firing function rate \\
	$u(\mathbf{r},t)$ & external input \\
	$\Omega$ & spatial domain \\
	$f_{max}$ & maximal firing rate \\
	$\varsigma$ & slope of sigmoidal activation function \\
	$v_0$ & firing threshold \\
	$\delta(t)$ & Dirac-delta function \\
	$L$ & temporal differential operator \\
	$e(\mathbf{r},t)$ & field disturbance, covariance $\gamma$\\
	$m(\mathbf{r},\mathbf{r}')$ & sensor kernel, variance $\sigma_m^2$ \\
	$\epsilon(\mathbf{r}_n,t)$ & observation noise, covariance $\Sigma_\epsilon$ \\
	$y(\mathbf{r}_n,t)$ & obseration \\
	$n$ & sensor index $n=1,..,N$ \\
	$T$ & time step \\
	$\xi$ & time constant parameter \\
	$\mathbf{\phi(r)}$ & vector of Gaussian basis functions \\
	$\mathbf{x}_t$ & state vector at time $t$ \\
	$\mathbf{\psi}$ & vector of connectivity kernel basis functions \\
	$\theta$ & vector of connectivity kernel parameters \\
	$\Gamma$ & inner product of field basis functions \\
	$q()$ & state function \\
	$k()$ & maps into to discrete state of input function \\
	$\mathbf{e}_t$ & state disturbance, covariance $\Sigma_e$ \\
	$\mathbf{C}$ & observation matrix \\
	$V(\nu)$ & Spectrum of the dynamic field \\ 
	$\nu$ & Spatial frequency \\
	$\nu_c$ & Spatial cut-off frequency \\
\end{tabular}

\begin{tabular}{c|c}
	\hline\hline
	Symbol & Description \\ \hline
	$\Delta_s$ & Distance between adjacent sensors \\
	$\rho$ & Over-sampling parameter \\	
	$\Delta_b$ & Distance between field basis functions \\	
	$\sigma_{nu}^2$ & Variance of FT of Gaussian basis function \\
	$\sigma^2$ & Spatial variance of Gaussian in spatial domain \\
	$\hat{\mathbf{x}}$ & State estimate \\
	$\chi$ & Matrix of sigma vectors \\
	$\hat{\mathbf{x}}_t^{f-}$ & Forward \emph{a priori} estimate \\
	$\hat{\mathbf{x}}_t^f$ & Forward \emph{a posteriori} estimate \\
	$\hat{\mathbf{x}}_t^{b-}$ & Backward \emph{a priori} estimate \\
	$\hat{\mathbf{x}}_t^{b}$ & Backward \emph{a posteriori} estimate \\
	$P^f_t$ & Forward \emph{a posteriori} covariance matrix \\
	$P^{f-}_t$ & Forward \emph{a priori} covariance matrix \\	
\end{tabular}
\section*{References}
\bibliographystyle{unsrt}
\bibliography{BrainIDE}
